Metal Mixing and Ejection in Dwarf Galaxies is Dependent on Nucleosynthetic Source

### 2018-07-31 main.tex
working on conclusions.

\subsection{Impact on Stellar Abundance Patterns}
\label{sec:stellar abundances}
In order to better understand the physics driving stellar abundance patterns and distributions it is important to characterize the chemical evolution of star-forming gas in our simulations. This could be used to help disentangle sources of scatter in observed stellar abundance, including radial/azimuthal abundance gradients and stellar migration, redshift evolution, asymmetric accretion of pristine gas, and the intrinsic scatter in ISM abundances. If the 
%mm star-forming gas in our simulations can also be well-fit by a log-normal + power-law PDF
        elemental abundances in star-forming gas in our simulations can also be well fit by 
        log-normal + power-law PDFs,
and if we can parameterize their evolution as functions of global galaxy properties, this %mm can 
    could
be used as a powerful 
%mm and analytic 
tool for modeling stellar abundance patterns in semi-analytic models. This would be a physically motivated way to account for both the intrinsic spread in stellar abundances due to inhomogeneities in the ISM and variations in mixing for different metal species. We reserve an analysis of these distributions in abundance space in both the gas and stars for future work. However, we briefly discuss the connection to star-forming gas and stellar enrichment below.

Unfortunately, there is insufficient star-forming gas at any one time in our simulations to construct PDFs of 
%mm metal distributions
    elemental abundances. 
However, star-forming gas originates from the CNM, so it is reasonable to expect that this gas, and therefore stars themselves, 
%mm follow a similar PDF 
   to have abundance PDFs similar
to the CNM. To verify this, Figure ~\ref{fig:stars} shows the difference between the oxygen mass fraction of stars and the median mass fraction from the CNM PDF at the time that particle formed. The large scatter at early times ($<$ 120~Myr) is a result of the early enrichment phase, when the initial gas oxygen mass fraction was zero. Stars seem to be sampled evenly around the median of the CNM distribution, with only a slight bias (52\% of stars) towards values below the median. However, for stars formed after the initial phase, the median separation from the CNM median is 0.31~dex, and can reach up to $\sim$~0.5~dex. Though a significant deviation, this is smaller than the typical inner-quartile range (IQR) of the CNM (see Figure~\ref{fig:phase-statistics}). Additionally, if we further subdivide the CNM by density, the 
%mm metal fraction 
    abundance
PDFs narrow and tend towards the median value as a function of increasing density. High density gas, from which star formation occurs, is not biased towards higher metallicity in cold gas.

@@

\subsection{Implications for Exotic Enrichment Sources}
\label{sec:exotic enrichment}
%   AE: Benoit suggested removing arguments relating to NS-NS merger energetics as I had misred / miscaclulated the energy release values. He said greater emphasise on potential differences of mixing due to rarity, rather than energy.
We have shown that the dynamical evolution of metals depends upon their nucleosynthetic source, focusing here on AGB synthesized elements and SN synthesized elements. However, these differences should also apply to exotic enrichment sources, such as hypernovae (HNe), neutron-star neutron-star mergers (NSM), and neutron star - black hole mergers. These sources can have energies that differ significantly from typical SN, reaching $> 10^{52}$~erg for HNe \citep{Nomoto2004}, for example. In addition, the rarity of these events, for example, there are approximately $10^{-5}$ per M$_{\odot}$ of star formation for NSMs \citep{Kim2015} in the Milky Way, could significantly influence the typical ISM environments within which they eject their metals. We would therefore expect different mixing behaviors for metals synthesized through these channels as compared to SNe. Based on our results here, we would expect elements from HNe to be more well mixed in the ISM of all galaxies, but more readily ejected in dwarf galaxies, as compared to elements from SNe. 
% Benoit suggesed removing this given energy argument removal above
%    Elements from NSMs, a significant source of r-process enrichment, would be less well-mixed in the ISM of all galaxies and more easily retained than SNe elements in dwarf galaxies. 
Depending on their injection energy, NSMs could exhibit different mixing behaviors in the ISM. These differences could provide important signatures for distinguishing individual, exotic enrichment sources from observed stellar abundance patterns in our own Milky Way and in nearby dwarf galaxies. In particular, low mass dwarf galaxies with unusual r-process enrichment \citep[e.g.][]{Ji2016}, are valuable for constraining the source of these elements, their frequency, and typical yields. The differential metal evolution presented in our work both opens up an additional avenue by which elements from distinct nucleosynthetic sources may be distinguished in observations and challenges current assumptions used in interpreting these observations.

@@

\section{Conclusions}
\label{sec:conclusions}
For the first time we present a detailed analysis of galactic chemodynamics and metal mixing on an element-by-element basis in a low mass dwarf galaxy with hydrodynamics simulations that simultaneously capture multi-channel stellar feedback in detail with a multi-phase ISM. This high resolution simulation, coupled with our star-by-star modelling of stellar nucleosynthetic yields, has allowed us to analyze how individual metal species couple to the galactic winds and ISM of this galaxy. We find that individual metal species do not share the same dynamical evolution, with differences directly related to nucleosynthetic origin (AGB winds, winds from massive stars, core collapse SNe, and Type Ia SNe). This difference is most significant in our model between elements ejected predominately through AGB winds and those ejected predominantly by core-collapse SNe. In addition, we find for the first time that the metal mass fraction PDFs of each metal in the ISM can be described using an analytic piece-wise log-normal + power-law PDF. The properties of these PDFs vary with both ISM phase and metal species, again driven mostly by differences in enrichment sources.

@@

\item Power-law tails in metal fraction PDFs are a natural consequence of ongoing chemical enrichment, with the power-law slope related to the rate at which mixing dilutes newly enriched gas.

>>>>

\subsection{Impact on Stellar Abundance Patterns}
\label{sec:stellar abundances}
In order to better understand the physics driving stellar abundance patterns and distributions it is important to characterize the chemical evolution of star-forming gas in our simulations. This could be used to help disentangle sources of scatter in observed stellar abundance, including radial/azimuthal abundance gradients and stellar migration, redshift evolution, asymmetric accretion of pristine gas, and the intrinsic scatter in ISM abundances. If the 
%mm star-forming gas in our simulations can also be well-fit by a log-normal + power-law PDF
        metal mass fractions in star-forming gas in our simulations 
        can also be well fit by log-normal + power-law PDFs,
and if we can parameterize their evolution as functions of global galaxy properties, this 
%mm can 
    could
be used as a powerful 
%mm and analytic 
tool for modeling stellar abundance patterns in semi-analytic models. This would be a physically motivated way to account for both the intrinsic spread in stellar abundances due to inhomogeneities in the ISM and variations in mixing for different metal species. We reserve an analysis of these distributions in abundance space in both the gas and stars for future work. However, we briefly discuss the connection to star-forming gas and stellar enrichment below.

Unfortunately, there is insufficient star-forming gas at any one time in our simulations to construct PDFs of 
%mm metal distributions
    metal mass fractions. 
However, star-forming gas originates from the CNM, so it is reasonable to expect that this gas, and therefore stars themselves, 
%mm follow a similar PDF 
   to have mass fraction PDFs similar
to the CNM. To verify this, Figure ~\ref{fig:stars} shows the difference between the oxygen mass fraction of stars and the median mass fraction from the CNM PDF at the time that particle formed. The large scatter at early times ($<$ 120~Myr) is a result of the early enrichment phase, when the initial gas oxygen mass fraction was zero. Stars seem to be sampled evenly around the median of the CNM distribution, with only a slight bias (52\% of stars) towards values below the median. However, for stars formed after the initial phase, the median separation from the CNM median is 0.31~dex, and can reach up to $\sim$~0.5~dex. Though a significant deviation, this is smaller than the typical inner-quartile range (IQR) of the CNM (see Figure~\ref{fig:phase-statistics}). Additionally, if we further subdivide the CNM by density, the metal fraction PDFs narrow and tend towards the median value as a function of increasing density. High density gas, from which star formation occurs, is not biased towards higher metallicity in cold gas.

@@

\subsection{Implications for Exotic Enrichment Sources}
\label{sec:exotic enrichment}
%   AE: Benoit suggested removing arguments relating to NS-NS merger energetics as I had misred / miscaclulated the energy release values. He said greater emphasise on potential differences of mixing due to rarity, rather than energy.
We have shown that the dynamical evolution of metals depends upon their nucleosynthetic source, focusing here on AGB synthesized elements and SN synthesized elements. However, these differences should also apply to exotic enrichment sources, such as hypernovae, neutron-star neutron-star mergers, and neutron-star black-hole mergers. These sources can have energies that differ significantly from typical SNe, reaching $> 10^{52}$~erg for hypernovae \citep{Nomoto2004}, for example. In addition, 
%mm the rarity of these events, for example,  there are approximately $10^{-5}$ per M$_{\odot}$ of star formation NSM in the Milky Way, could
     these events are rare.  For example, neutron star mergers occur 
     at a rate of approximately $10^{-5}$ per M$_{\odot}^{-1}$ of star
     formation in the Milky Way\citep{Kim2015}.  This rarity could 
significantly influence the typical ISM environments into which they eject their metals. We would therefore expect different mixing behaviors for metals synthesized through these channels as compared to SNe. Based on our results here, 
%mm
     for example,
we would expect elements from hypernovae to be more well mixed in the ISM of all galaxies, but more readily ejected in dwarf galaxies, as compared to elements from SNe. 
% Benoit suggesed removing this given energy argument removal above
%    Elements from NSMs, a significant source of r-process enrichment, would be less well-mixed in the ISM of all galaxies and more easily retained than SNe elements in dwarf galaxies. 
Depending on their injection energy, neutron star mergers could exhibit different mixing behaviors in the ISM. These differences could provide important signatures for distinguishing individual, exotic enrichment sources from observed stellar abundance patterns in our own Milky Way and in nearby dwarf galaxies. In particular, low mass dwarf galaxies with unusual r-process enrichment \citep[e.g.][]{Ji2016}, are valuable for constraining the source of these elements, their frequency, and typical yields. The differential metal evolution presented in our work both opens up an additional avenue by which elements from distinct nucleosynthetic sources may be distinguished in observations and challenges current assumptions used in interpreting these observations.

@@

\section{Conclusions}
\label{sec:conclusions}
%mm For the first time we 
We present a detailed analysis of galactic chemodynamics and metal mixing on an element-by-element basis in a low mass dwarf galaxy with hydrodynamics simulations that simultaneously capture multi-channel stellar feedback in detail with a multi-phase ISM. This high resolution simulation, coupled with our star-by-star modelling of stellar nucleosynthetic yields, has allowed us to analyze
%mm
      for the first time
how individual metal species couple to the ISM and galactic wind of this galaxy. We find that individual metal species do not share the same dynamical evolution, with differences directly related to nucleosynthetic origin (AGB winds, winds from massive stars, or core collapse SNe)
%mm , and Type Ia SNe).  [I'm inclined to leave them out as we really didn't model their influence]
This difference is most significant in our model between elements ejected predominately by AGB winds and those ejected predominantly by core-collapse SNe. In addition, we find 
%mm for the first time 
    the novel result
that the mass fraction PDFs of each metal in the ISM can be described using an analytic piecewise log-normal + power-law PDF. The properties of these PDFs vary with both ISM phase and metal species, again driven mostly by differences in enrichment sources.

@@

\item Power-law tails 
%mm in 
   on log-normal 
metal fraction PDFs are a natural consequence of ongoing chemical enrichment, with the power-law slope related to the rate at which mixing dilutes newly enriched gas.
